### 2018-05-04 gf-icfp-2018/RESPONSE-ICFP-2018.txt

ICFP '18 Paper #21 Author Response
===========================================================================

Review #21B mischaracterizes our results in at least three ways.

FIRST, the review misrepresents our theoretical results as variations of
known theorems.

This claim is _completely wrong_ (with an exception). Our theorems expose
the looseness of previous claims in the literature concerning "soundness"
in the area of gradual typing. Our meta-theorems (and their logical
implications) explain that these claims are incomparable and must be read
very carefully. Technically our meta-theorems show that soundness results
can be stated only in terms of properties preserved through run-time
configurations. As a result, the implications for the programmer differ
dramatically. For details, see below.

SECOND, the review #21B seems to imply that our results might be preempted
by two papers (one of which is still unavailable in public):

@@

> _A Framework for Object-Oriented Gradual Typing_.
>  Chung, Li, Zappa Nardelli, Vitek.
>  ECCOP 2018
>
> _Efficient Gradual Typing_.
> Kuhlenschmidt, Almahallawi, Siek.
> https://arxiv.org/abs/1802.06375

Felleisen is the PI on a large NSF grant that also supports these efforts.
The papers were presented at the August 2017 PI day at Northeastern. When
we submitted our paper, Chung et al.'s work was NOT ACCEPTED and NOT
AVAILABLE in public. [[ We confirmed this statement with the authors who
sit across the hallway from us. ]]  Now that its status has changed we can
easily compare the two research results and will do so in the paper. For
a detailed comparison, see below.

We do not consider Kuhlenschmidt et al.'s extremely preliminary work
related and we do not accept that it contradicts our result. In brief,
Kuhlenschmidt et al. are working on an alternative approach to the
compilation of NEWLY DESIGNED gradually typed languages. Their work has
confirmed that on a very small number of micro-benchmarks, the compiler can
eliminate a large fraction of overhead due to the TRANSIENT semantics of
gradual semantics.

Kuhlenschmidt et al.'s work does not include the apples-to-apples
comparison of work, with full-fledged implementations of all three major
semantics. Similarly, while our three implementations can deal with a dozen
non-trivial benchmarks (programs used on a daily basis), Kuhlenschmidt et
al.'s compiler cannot deal with anything realistic so far. Of course, we do
look forward to the day when their work supersedes the current state of the
art with respect to performance.  We wish to use gradually typed languages
on a daily basis.

THIRD, the review claims that our paper does not support our statements
about performance. Specifically, the review claims

> p19. "Sound static types can eliminate the need to dispatch, thus the
> natural embedding's performance can exceed that of the erasure
> embedding"
> This paper does not provide empirical results regarding this claim.

The reviewer seems to have overlooked TABLE ____ttt___ on PAGE ___ppp___.
This table clearly demonstrate that the performance of NATURAL dominates
the performance of LOCALLY-DEFENSIVE on programs whose main modules are
fully typed (though they may use untyped run-time libraries).

@@

Response to Review #21A
===========================================================================

> A weak part of the paper---which is easy to fix---is that there is no
> discussion of mixed-semantics languages.  As such, Dart is badly characterized
> in this paper as an optional/erasure style language.

Thank you for reminding us of Dart's mixed approach. We will include a
brief discussion and draw attention to its approach, also as an area for
future work.

@@

> - In Theorem 2.0, 2nd reduction alternative, I don't understand why the right
>   hand side cannot be simply: E[dyn t' E* [TagErr]], considering that e' steps
>   to TarErr and that the reduction relation admits reduction under E*.
>   (same question for Theorem 2.6)

You are correct. Thanks!

> - The data of Figure 12 could be plotted (again, there is space)

Yes! This table is too easy to miss.

This is a correct observation. The Indiana Typed Racket team is revising
the Pycket compiler for Racket to address this issue. We are looking
forward to reproducing our experimental work on top of that new
backend. But this area remains future work for sound gradually typed
approaches.

Detailed Response to Review #21B
===========================================================================

### 2018-05-04 gf-icfp-2018/RESPONSE-ICFP-2018.txt

ICFP '18 Paper #21 Author Response
===========================================================================

Review #21B mischaracterizes our results in at least three ways.

FIRST, the review misrepresents our theoretical results as variations of
known theorems.

This claim is _wrong_. Our theorems articulate statements from the existing
literature in a single framework, making them comparable. [[ We clearly
must explicate how each section relates to prior literature directly. ]] On
this apples-to-apples platform, our theorem statements expose the looseness
of previous statements in the literature concerning "soundness" in the area
of gradual typing. These three pairs of theorems explain that the
statements in the literature denote distinct notions of soundness and must
be read very carefully. For technical details, see below.

@@

SECOND, the review #21B seems to imply that our results might be preempted
by two papers (one of which is still unavailable in public):

> _A Framework for Object-Oriented Gradual Typing_.
>  Chung, Li, Zappa Nardelli, Vitek.
>  ECOOP 2018
>
> _Efficient Gradual Typing_.
> Kuhlenschmidt, Almahallawi, Siek.
> https://arxiv.org/abs/1802.06375

Felleisen is the PI on a large NSF grant that also supports these efforts.
The papers were presented at the August 2017 PI day at Northeastern. When
we submitted our paper, Chung et al.'s work was NOT ACCEPTED and NOT
AVAILABLE in public. [[ We confirmed this statement with the authors who
sit across the hallway from us. ]]  Now that its status has changed we can
easily compare the two research results and will do so in the paper. For
a detailed comparison, see below.

We consider Kuhlenschmidt et al.'s very-extremely preliminary work.  In
brief, Kuhlenschmidt et al. are working on an alternative approach to the
compilation of NEWLY DESIGNED gradually typed languages. The language is
extremely small and the paper warns readers not to use the compiler
"because it does not support a full language" (p. 10, left).  Their work
confirms that on a small number of micro-benchmarks, their Grift compiler
performs differently for four different styles of casts.

But, we do not accept that this status report preempts any of our practical
results. Kuhlenschmidt et al.'s work does not include an apples-to-apples
comparison with full-fledged implementations of the three major
semantics. Similarly, while our three implementations can deal with
non-trivial benchmarks (programs used on a daily basis), Kuhlenschmidt et
al.'s compiler cannot deal with anything realistic.

@@

THIRD, the review claims that our paper does not support our statements
about performance. Specifically, the review claims

> p19. "Sound static types can eliminate the need to dispatch, thus the
> natural embedding's performance can exceed that of the erasure
> embedding"
> This paper does not provide empirical results regarding this claim.

FIGURE 12 on PAGE ___ppp___ LINE __lll___ clearly demonstrates that the
performance of NATURAL dominates the performance of ERASURE on programs
when the typed-driven optimizer can exploit sound information.

@@

Response to Review #21C
===========================================================================

> I am a little bit surprised at the fact that static typing does not
> contribute the runtime performance at all.  I would like to see the
> runtime performance on a JIT compiler or a compiler equipped with
> type-conscious optimization.

This is a correct observation. The Indiana Typed Racket team is revising
the Pycket compiler for Racket to address this issue. We are looking
forward to reproducing our experimental work on top of that new
backend. But this area remains future work for sound gradually typed
approaches.

Response to Review #21B
===========================================================================

@@

> + The paper does a thorough and controlled performance experiment that
>   confirms that the results of Vitousek et al. (2017) on Reticulated
>   Python carry over to the setting of Typed Racket. That is, the
>   transient semantics avoids the catastrophic performance problems
>   that occur in naive implementations of the natural approach, but
>   transient incurs overhead in statically typed code.

Vitousek et al. (2017) does not directly compare transient to natural.

Remarks like this are exactly why we prepared this paper. It is time to stop
making unscientific claims about different languages with different type
systems and different guarantees.

@@

> * The semantic framework is a straightforward adaptation of
>   the Matthews-Findler multi-language semantics.

The first sentence of Section 2.2 and first paragraph of Section 6 acknowledge
Matthews & Findler.

@@

Correct, we will discuss the relationship.

> * Theorem 2.6 is an adaptation of Corollary 5.5.1 (Type soundness) of
>   Vitousek et al. (2017).  The paper does not discuss this close
>   relationship.

Theorem 2.6 is **not** an adaptation of Vitousek et al.'s "type soundness".
Vitousek et al.'s Corollary 5.5.1 is about a single language.

Nor is our Theorem 2.6 an adaptation of Vitousek et al.'s Theorem 5.5
(Open-World Soundness):

- Vitousek et al. do not define a type system for the surface language;
  it is impossible to reason about programs independent of their core-language
  translation (via Vitousek et al.'s $\rightsquigarrow$).

- Vitousek et al.'s "stuck terms" are defined in terms of the core language;
  it is not clear from their theorem that well-typed expressions cannot get
  stuck. Again, one must study the translation of surface-language expressions
  into the core language. In contrast, our theorems clearly state the
  different guarantees for statically-typed and dynamically-typed expressions.

@@

> p1. The abstract is much too short.

If you believe something is missing from the abstract, we will consider
including it.

> p1. "But, it may impose a huge run-time cost"
> It would be appropriate to cite recent work by Kuhlenschmidt et
> al. (see above) that shows that the natural semantics does not have to
> impose a huge run-time cost.

We disagree, it would not be appropriate to cite Kuhlenschidt et al. here.
The sentence above is trying to communicate the historical fact that the
performance of the natural embedding led researchers and implementors to
consider other alternatives. Kuhlendschmidt et al.'s work does not change the
past.

@@

> p1. In the discussion of "type soundness", this paper should discuss
> the Blame Theorem, which provides the same guarantees that are provided
> by the theorems in this paper.

None of the blame theorems (that we know) provide the same guarantees.

The blame theorems in Wadler and Findler (2009), Ahmed et al. (2011), and
Vitousek et al. (2017) are guarantees about casts in an intermediate language.
Our soundness theorems are guarantees about the results of evaluating a
surface-language term. These are different guarantees.

@@

> p9. "Second, monitoring adds prohibitive run-time cost."
> That is only true of naive implementations. Again, see
> Kuhlenschmidt et al.

Kuhlenschmidt et al's work is not relevant here --- the sentence above is
summarizing the motivation for transient given in Vitousek et al. (2017).

### 2018-05-04 gf-icfp-2018/RESPONSE-ICFP-2018.txt

This claim is _wrong_. Our theorems articulate statements from the existing
literature in a single framework, making them comparable. [[ We clearly
must explicate how each section relates to prior literature directly. ]] On
this apples-to-apples platform, our theorem statements expose the looseness
of previous statements in the literature concerning "soundness" in the area
of gradual typing. These three pairs of theorems explain that the
statements in the literature denote distinct notions of soundness and must
be read very carefully. For technical details, see below.

SECOND, the review #21B seems to imply that our results might be preempted
by two papers (one of which is still unavailable in public):

> _A Framework for Object-Oriented Gradual Typing_.
>  Chung, Li, Zappa Nardelli, Vitek.
>  ECOOP 2018
>
> _Efficient Gradual Typing_.
> Kuhlenschmidt, Almahallawi, Siek.
> https://arxiv.org/abs/1802.06375

Felleisen is the PI on a large NSF grant that also supports these efforts.
The papers were presented at the August 2017 PI day at Northeastern. When
we submitted our paper, Chung et al.'s work was NOT ACCEPTED and NOT
AVAILABLE in public. [[ We confirmed this statement with the authors who
sit across the hallway from us. ]]  Once its status is changed, we will
compare the two research results in the paper.

Kuhlenschmidt et al. are working on an alternative approach to the compilation
of NEWLY DESIGNED gradually typed languages. The language is a work-in-progress;
indeed, the paper warns readers not to use the compiler "because it does not
support a full language" (p. 10, left). Their work compares two kinds of
references and two kinds of casts on an ad-hoc sample of mixed-typed
configurations derived from 6 micro-benchmarks.

But, we do not accept that this status report preempts any of our practical
results. Kuhlenschmidt et al.'s work does not include an apples-to-apples
comparison with full-fledged implementations of the three major semantics.
Similarly, while our evaluation measures non-trivial benchmarks,
Kuhlenschmidt et al.'s compiler cannot handle most of the shootout benchmarks.

@@

THIRD, the review claims that our paper does not support our statements
about performance. Specifically, the review claims

> p19. "Sound static types can eliminate the need to dispatch, thus the
> natural embedding's performance can exceed that of the erasure
> embedding"
> This paper does not provide empirical results regarding this claim.

FIGURE 12 on PAGE 15 (LINE 730) clearly demonstrates that the
performance of NATURAL dominates the performance of ERASURE on programs
when the typed-driven optimizer can exploit sound information.

@@

Vitousek et al. (2017) does not directly compare transient to natural.

Remarks like this are exactly why we prepared this paper. It is time to stop
making unscientific claims about different languages with different type
systems and different guarantees.

> * The semantic framework is a straightforward adaptation of
>   the Matthews-Findler multi-language semantics.

The first sentence of Section 2.2 and first paragraph of Section 6 acknowledge
Matthews and Findler.

@@

> * Theorem 2.0 is a just a combination of the usual Type Safety theorem
>   for gradually typed languages (Siek and Taha 2006) and the Blame
>   Theorem (Tobin-Hochstadt and Felleisen 2006, Wadler and Findler
>   2009). The paper does not discuss these relationships.

Incorrect. The usual type soundness theorems relate ONE typing property to a
semantics. Theorem 2.0 relates a surface-language typing property to a DIFFERENT
evaluation-language property and a semantics. [[ Critically, surface-language
typing is not preserved by the semantics. ]]

To clarify our point from the general response above, theorems 2.0, 2.3, and 2.6
demonstrate how the various "type soundness" theorems from the literature are
superficially similar, but rely on fundamentally different evaluation- language
typing properties.

We will try to clarify these important subtleties.

@@

> * Theorems 2.1, 2.3, and 2.4 are all the usual type safety theorem for
>   dynamically typed languages.

Incorrect. The "usual" type safety theorem relates ONE safety property to
a semantics. In contrast, theorems 2.1, 2.3, and 2.4 relate a surface-language
safety property to an evaluation-language safety property and semantics.
An evaluation-language property is NECESSARY to handle the run-time interactions
between typed and untyped code, and its meaning DEPENDS on the desired notion of
type soundness.

@@

> * Theorem 2.5 is just type safety for a statically typed program, and
>   is a corollary of the equivalence between gradual typing and static
>   typing on fully annotated programs (Siek et al. 2015).  The paper
>   does not discuss this close relationship.

Correct, we will discuss the relationship.

> * Theorem 2.6 is an adaptation of Corollary 5.5.1 (Type soundness) of
>   Vitousek et al. (2017).  The paper does not discuss this close
>   relationship.

Theorem 2.6 is NOT an adaptation of Vitousek et al.'s "type soundness".
Vitousek et al.'s Corollary 5.5.1 is about a single language.

Nor is our Theorem 2.6 a simple adaptation of Vitousek et al.'s Theorem 5.5
(Open-World Soundness):

- Vitousek et al. do not define a type system for the surface language;
  it is impossible to reason about programs in their system independent of the
  translation from surface to core (via Vitousek et al.'s $\rightsquigarrow$).

@@

Kuhlenschmidt et al's work is not relevant here --- the sentence above is
summarizing the motivation for transient given in Vitousek et al. (2017).

Please let us know if you believe we have mis-represented Vitousek et al. (2017)

> p4. It is confusing to use boundary errors for both cast errors and
> for errors such as div-by-zero.

Cast errors and div-by-zero errors are the same kind of error. They
both arise when one language sends invalid data to another.

In other words, soundness for a pair of languages is a natural generalization
of soundness for a single language that includes primitive operations
(such as division) in its trusted code base.

@@

> p17. "the implicit conjectures of the literature"
> You seem to be overlooking the performance experiments
> reported on by Vitousek et al. (2017).

Vitousek et al. (2017) do not measure the performance of programs that mix
typed and untyped code.

> p19. "The natural embedding incurs three significantly larger kinds of
> costs." You're forgetting the fourth: the natural embedding also has
> the space-efficiency problem pointed out by Herman et al. (2006),
> which also induces problems with time efficiency.

We did not forget, but you are right we need to be clearer that the cost
of ALLOCATION is a space+time cost and is unbounded in principle.

### 2018-05-04 gf-icfp-2018/RESPONSE-ICFP-2018.txt

ICFP '18 Paper #21 Author Response
===========================================================================

Review #21B mischaracterizes our results in at least three ways.

FIRST, the review misrepresents our theoretical results as variations of
known theorems.

This claim is _wrong_. Our theorems articulate statements from the existing
literature in a single framework, making them comparable. [[ We clearly
must explicate how each section relates to prior literature directly. ]] On
this apples-to-apples platform, our theorem statements expose the looseness
of previous statements in the literature concerning "soundness" in the area
of gradual typing. Our three pairs of theorems explain that the statements in
the literature denote distinct notions of soundness and must be read very
carefully. For technical details, see below.

@@

The first sentence of Section 2.2 and first paragraph of Section 6 acknowledge
Matthews and Findler.

> * Theorem 2.0 is a just a combination of the usual Type Safety theorem
>   for gradually typed languages (Siek and Taha 2006) and the Blame
>   Theorem (Tobin-Hochstadt and Felleisen 2006, Wadler and Findler
>   2009). The paper does not discuss these relationships.

Incorrect. The usual type soundness theorems relate ONE typing property to a
semantics. Theorem 2.0 relates a surface-language typing property to a DIFFERENT
evaluation-language property and a semantics. [[ Critically, surface-language
typing is not preserved by the semantics. ]]

To clarify our point from the general response above, theorems 2.0, 2.3, and 2.6
demonstrate how the various "type soundness" theorems from the literature are
superficially similar, but rely on fundamentally different evaluation-language
typing properties.

We will explain these important subtleties in the paper.

@@

> p1. The abstract is much too short.

Why?

> p1. "But, it may impose a huge run-time cost"
> It would be appropriate to cite recent work by Kuhlenschmidt et
> al. (see above) that shows that the natural semantics does not have to
> impose a huge run-time cost.

We disagree, it would not be appropriate to cite Kuhlenschmidt et al. here.
The sentence above is trying to communicate the historical fact that the
performance of the natural embedding led researchers and implementors to
consider other alternatives. Kuhlenschmidt et al.'s work does not change the
past.

@@

> p19. "The natural embedding incurs three significantly larger kinds of
> costs." You're forgetting the fourth: the natural embedding also has
> the space-efficiency problem pointed out by Herman et al. (2006),
> which also induces problems with time efficiency.

We did not forget, but you are right we need to be clearer that the cost
of ALLOCATION is a space+time cost and is unbounded in principle.

We will clarify.

> The paper should describe its relationship with [Chung et al. 2018]

Chung et al. (2018) define a calculus for object-oriented gradual typing and
present four translations from a common surface language into the calculus.
Through various examples, they demonstrate how the different translations can
lead to core-calculus programs with different semantics. In contrast to our
work, Chung et al. do not prove type soundness in terms of the surface language
and do not evaluate the performance of their translations.

### 2018-05-04 gf-icfp-2018/RESPONSE-ICFP-2018.txt

This claim is _wrong_. Our theorems articulate statements from the existing
literature in a single framework, making them comparable. [[ We clearly
must explicate how each section relates to prior literature directly. ]] On
this apples-to-apples platform, our theorem statements expose the looseness
of previous statements in the literature concerning "soundness" in the area
of gradual typing. Our three pairs of theorems explain that the statements in
the literature denote distinct notions of soundness and must be read very
carefully. For technical details, see below.

SECOND, the review #21B seems to imply that our results might be preempted
by two papers (one of which is still unavailable in public):

> _A Framework for Object-Oriented Gradual Typing_.
>  Chung, Li, Zappa Nardelli, Vitek.
>  ECOOP 2018
>
> _Efficient Gradual Typing_.
> Kuhlenschmidt, Almahallawi, Siek.
> https://arxiv.org/abs/1802.06375

Felleisen is the PI on a large NSF grant that also supports these efforts.
The papers were presented at the August 2017 PI day at Northeastern. When
we submitted our paper, Chung et al.'s work was NOT ACCEPTED and NOT
AVAILABLE in public. [[ We confirmed this statement with the authors who
sit across the hallway from us. ]]  Once its status is changed, we will
compare the two research results in the paper. See the end of the response
for a proposed prose addition to our paper, because this is the closest
piece of research in existence.

@@

We did not forget, but you are right we need to be clearer that the cost
of ALLOCATION is a space+time cost and is unbounded in principle.

We will clarify.

> The paper should describe its relationship with [Chung et al. 2018]

Here is what we propose to add to the paper:

```
Chung et al. (2018) study the relationship of four different designs of
gradually typed object-oriented syntaxes. To this end, they develop a core
language, dubbed Kafka, and prove a type-soundness theorem for it.  They
define four translations from the four syntaxes to Kafka. For the
comparative part of their work, they use examples to demonstrate how the
four systems (syntax plus translation) can lead to seemingly similar
core-calculus programs with rather different behavior.

By contrast, this paper assigns three different semantics to the same
syntax and type system. It then articulates precise soundness theorems
about each of these semantics, which give rise to a natural comparison of
the subtleties underlying the known variants of gradually typed
implementations. Also, Chung et al. do not implement their systems and do
not present comparative performance evaluation results in their framework. 
'''

### 2018-05-04 gf-icfp-2018/RESPONSE-ICFP-2018.txt

ICFP '18 Paper #21 Author Response
===========================================================================

Review #21B mischaracterizes our results in three major ways.

@@

FIRST, the review misrepresents our theoretical results as variations of
known theorems.

This claim is _wrong_. Our theorems articulate statements from the existing
literature in a single framework, making them comparable. [[ We clearly
must explicate how each section captures prior literature directly. ]] On
this apples-to-apples platform, our theorem statements expose the looseness
of previous statements in the literature concerning "soundness" in the area
of gradual typing. The three pairs of theorems explain that the statements in
the literature denote distinct notions of soundness and must be read very
carefully. For technical details, see below.

SECOND, the review seems to imply that parts of our results might be preempted
by two papers:

@@

> _A Framework for Object-Oriented Gradual Typing_.
>  Chung, Li, Zappa Nardelli, Vitek.
>  ECOOP 2018
>
> _Efficient Gradual Typing_.
> Kuhlenschmidt, Almahallawi, Siek.
> https://arxiv.org/abs/1802.06375

Felleisen is the PI on a large NSF grant that also supports these efforts.
The papers were presented at the August 2017 PI day at Northeastern. When
we submitted our paper, Chung et al.'s work was NOT ACCEPTED and NOT
AVAILABLE in public. [[ We confirmed this statement with the authors who
sit across the hallway from us. ]]  Once its status is changed, we will
compare the two research results in the paper. See the end of the response
for a proposed prose addition to our paper, because this is the closest
piece of research in existence.

Kuhlenschmidt et al. are working on an alternative approach to the compilation
of a NEWLY DESIGNED gradually typed language. The language is a
work-in-progress; indeed, the paper warns readers not to use the compiler
"because it does not support a full language" (p. 10, left). Their work
compares two kinds of references and two kinds of casts on an ad-hoc sample of
mixed-typed configurations derived from 6 micro-benchmarks.

@@

This is a correct observation. The Indiana Typed Racket team is revising
the Pycket compiler for Racket to address this issue. We are looking
forward to reproducing our experimental work on top of that new
backend. But this area remains future work.

Response to Review #21B
===========================================================================

> + The paper does a thorough and controlled performance experiment that
>   confirms that the results of Vitousek et al. (2017) on Reticulated
>   Python carry over to the setting of Typed Racket. That is, the
>   transient semantics avoids the catastrophic performance problems
>   that occur in naive implementations of the natural approach, but
>   transient incurs overhead in statically typed code.

Vitousek et al. (2017) do not compare transient to natural.

Remarks like this are exactly why we conducted this research. It is time to stop
unscientific claims about different languages with different type
systems and different guarantees.

@@

> * The semantic framework is a straightforward adaptation of
>   the Matthews-Findler multi-language semantics.

The first sentence of Section 2.2 and first paragraph of Section 6 acknowledge
our debt to Matthews and Findler.

@@

> * Theorem 2.0 is a just a combination of the usual Type Safety theorem
>   for gradually typed languages (Siek and Taha 2006) and the Blame
>   Theorem (Tobin-Hochstadt and Felleisen 2006, Wadler and Findler
>   2009). The paper does not discuss these relationships.

Incorrect. The usual type soundness theorems relate ONE typing property to a
semantics. Theorem 2.0 relates a surface-language typing property to a DIFFERENT
evaluation-language property and a semantics. [[ Critically, surface-language
typing is not preserved by the semantics. ]]

To clarify our point from the general response above, theorems 2.0, 2.3, and 2.6
demonstrate how the various "type soundness" theorems from the literature are
superficially similar, but rely on fundamentally different evaluation-language
typing properties.

We will explicitly point out these important subtleties in the final version.

@@

> * Theorems 2.1, 2.3, and 2.4 are all the usual type safety theorem for
>   dynamically typed languages.

Incorrect. The "usual" type safety theorem relates ONE safety property to
a semantics. In contrast, theorems 2.1 and 2.4 relate a surface-language
safety property to an evaluation-language safety property and semantics.
An evaluation-language property is NECESSARY to handle the run-time interactions
between typed and untyped code, and its meaning DEPENDS on the desired notion of
type soundness.

Theorem 2.3 is completely different; it is a statement about well-typed
programs. This weak "type soundness" theorem explains the weak guarantees
provided by optionally-typed languages.

@@

> * Theorem 2.5 is just type safety for a statically typed program, and
>   is a corollary of the equivalence between gradual typing and static
>   typing on fully annotated programs (Siek et al. 2015).  The paper
>   does not discuss this close relationship.

Correct, we will explain the relationship in the final version.

@@

We will add a clarification to the paper.

> * Theorem 2.7 should be another example of the usual type safety
>   theorem for a dynamically typed language, but it looks like there
>   is a typo in saying |-_{LD} v : \lfloor \tau \rfloor
>   because there is no \tau to refer to here.

Thank you, that is definitely a typo.

Our comment above regarding theorems 2.1, 2.2, and 2.3 also holds for
theorem 2.7. It is an unusual theorem because it connects the surface language
to the evaluation language and uses two property judgments. The second
property judgment makes this theorem fundamentally different from 2.1 and 2.3.

@@

> p1. "But, it may impose a huge run-time cost"
> It would be appropriate to cite recent work by Kuhlenschmidt et
> al. (see above) that shows that the natural semantics does not have to
> impose a huge run-time cost.

We disagree, it would not be appropriate to cite Kuhlenschmidt et al. here.
The sentence above is trying to communicate the historical fact that the
performance of the natural embedding led researchers and implementors to
consider alternatives. Kuhlenschmidt et al.'s work does not change this fact.

@@

> p1. In the discussion of "type soundness", this paper should discuss
> the Blame Theorem, which provides the same guarantees that are provided
> by the theorems in this paper.

None of the blame theorems (that we know) provide the same guarantees.

The blame theorems in Wadler and Findler (2009), Ahmed et al. (2011), and
Vitousek et al. (2017) are guarantees about casts in an intermediate language.
Our soundness theorems are guarantees about the results of evaluating a
surface-language term. These are different guarantees.

But we must definitely relate our statement to these theorems.

@@

> p9. "Second, monitoring adds prohibitive run-time cost."
> That is only true of naive implementations. Again, see
> Kuhlenschmidt et al.

Kuhlenschmidt et al's work is not relevant here --- the sentence above is
summarizing the motivation for transient given in Vitousek et al. (2017).

> p4. It is confusing to use boundary errors for both cast errors and
> for errors such as div-by-zero.

Cast errors and div-by-zero errors are the same kind of error. They
both arise when one language sends invalid data to another. A div-by-zero
error comes from the run-time system. In many languages, the run-time system
relies on a low-level, untyped language.
